---
---

These bits are for mixing Nvim-R and renv. Nvim-R has a tool called nvimcom that runs the interaction and allows sending of things between the file and terminal buffers. It's packaged with Nvim-R, so much be installed. These instructions come from <https://gist.github.com/verasls/fc6596ee83dc820a6778d0e5a57f3d5c>, which I've modified as I use Packer for managing my neovim packages.

```{r renv_nvimcom_install, eval = FALSE}
renv::settings$ignored.packages("nvimcom")
install.packages("devtools")
devtools::install("~/.local/share/nvim/site/pack/packer/start/Nvim-R/R/nvimcom")
install.packages("languageserver")
```


```{r}
library(broom)
library(cowplot)
library(geojsonio)
library(gganimate)
library(ggthemes)
library(mapproj)
library(sf)
library(sugarbag)
library(tidyverse)
library(viridis)
```

```{r}
dat <-
  read_csv("county-area-list2011.csv")
```

```{r coronavirus_data}
coronavirus_data <-
  read_csv("https://api.coronavirus.data.gov.uk/v2/data?areaType=utla&metric=newCasesBySpecimenDate&metric=newAdmissions&format=csv")

coronavirus_data %>%
  ggplot(aes(x = date, y = newCasesBySpecimenDate, colour = areaCode)) +
  geom_line(show.legend = FALSE) +
  facet_wrap(vars(areaCode)) +
  theme_void() +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
  )
```

These data are outlines of UK [local administration units](https://geoportal.statistics.gov.uk/datasets/ons::local-administrative-units-level-1-january-2018-full-clipped-boundaries-in-united-kingdom?geometry=-61.936%2C46.017%2C57.068%2C63.434 "UK Office for National Statistics Local Administration Units Level 1 (2018)")]

```{r lau_areas}
lad_geo <-
  geojson_sf("https://opendata.arcgis.com/datasets/658297aefddf49e49dcd5fbfc647769e_4.geojson")

coronavirus_data %>%
  filter(date == "2021-04-01") %>%
  mutate(areaCode = toupper(areaCode)) %>%
  left_join(lad_geo, ., by = c("ctyua17cd" = "areaCode")) %>%
  ggplot() +
    geom_sf(aes(fill = newCasesBySpecimenDate), show.legend = "top") +
    theme_map()

covid_joined <-
  coronavirus_data %>%
    filter(date == "2021-04-01") %>%
    left_join(lad_geo, ., by = c("ctyua17cd" = "areaCode"))
```

```{r}
hex_size <- 0.6

utla_centroids <-
  lad_geo %>%
  create_centroids(sf_id = "ctyua17cd")
utla_grid <-
  utla_centroids %>%
  create_grid(
    hex_size = hex_size,
    buffer_dist = 12
  )
utla_hex_allo <-
  allocate(
    utla_centroids,
    hex_grid = utla_grid,
    sf_id = "ctyua17cd",
    hex_size = hex_size,
    hex_filter = 20,
    width = 30,
    verbose = TRUE
  )
hexes <-
  utla_hex_allo %>%
    fortify_hexagon(
      hex_size = hex_size,
      sf_id = "ctyua17cd"
    ) %>%
    left_join(
      coronavirus_data %>%
        filter(date == "2021-04-01"),
      by = c("ctyua17cd" = "areaCode")
    ) %>%
    mutate(
      poly_type = "hex",
      group = ctyua17cd
    )
```

```{r}
plot_map_figure <-
  function(
    shape_data,
    hexagon_data,
    fill_variable = "newCasesBySpecimenDate"
  ) {
  # Plot a figure of Local Authority data with both a choropleth map and tesselated hexagram.
  choropleth_map <-
    shape_data %>%
      ggplot() +
        geom_sf(
          aes_(fill = as.name(fill_variable)),
          lwd = 0.2
        ) +
        theme_map() +
        scale_fill_viridis(discrete = FALSE) +
        title("Map")
  choropleth_hexagram <-
    hexagon_data %>%
    ggplot(aes_(fill = as.name(fill_variable))) +
      geom_polygon(
        aes(x = long, y = lat, group = group)
      ) +
      theme_map() +
      scale_fill_viridis(discrete = FALSE) +
      title("Hexagram – each authority equally-sized")
  legend <-
    get_legend(
      choropleth_map +
        theme(legend.box.margin = margin(0, 0, 0, 15)) +
        labs(fill = "")
    )
  maps <-
    plot_grid(
      choropleth_map +
        theme(legend.position = "none"),
      choropleth_hexagram +
        theme(legend.position = "none"),
      labels = "AUTO",
      rel_widths = c(1.5, 2)
    )
  plot_grid(maps, legend, rel_widths = c(3, 0.6))
}

cp_title <-
  function(title) {
  ggdraw() +
    draw_label(
      title,
      fontface = "bold",
      x = 0,
      hjust = 0
    ) +
    theme(
      # add margin on the left of the drawing canvas,
      # so title is aligned with left edge of first plot
      plot.margin = margin(0, 0, 0, 7)
    )
  }
```

```{r}
map_w_legend <-
  plot_map_figure(
    shape_data = covid_joined,
    hexagon_data = hexes,
    fill_variable = "newCasesBySpecimenDate"
  )

title <-
  cp_title(
    "New cases of COVID-19 in UK Local Authorities on 1 April 2021."
  )

figure <-
  plot_grid(
    title,
    map_w_legend,
    ncol = 1,
    rel_heights = c(0.1, 1)
  )
save_plot(
  "map_figure.pdf",
  figure,
  base_height = 120,
  base_asp = 1.618,
  units = "mm"
)
```


```{r}
weekly_changes <-
  coronavirus_data %>%
    arrange(date) %>%
    group_by(areaCode) %>%
    mutate(
      weekly_case_change =
        newCasesBySpecimenDate - lag(newCasesBySpecimenDate, n = 7L),
      rel_changes =
        weekly_case_change / max(abs(weekly_case_change), na.rm = TRUE)
    )

weekly_case_change <-
  weekly_changes %>%
  ggplot(
    aes(
      x = date,
      y = rel_changes,
      colour = areaCode
    )
  ) +
    geom_line() +
    theme_void() +
    theme(
      legend.position = "none",
      strip.background = element_blank(),
      strip.text.x = element_blank()
    ) +
    facet_wrap(vars(areaCode))

save_plot(
  "time_figure.pdf",
  plot = weekly_case_change,
  base_height = 120,
  base_asp = 1.618,
  units = "mm"
)
```

```{r}
weekly_change_map <-
  weekly_changes %>%
    filter(date == "2021-04-01") %>%
    left_join(lad_geo, ., by = c("ctyua17cd" = "areaCode"))
weekly_change_hexes <-
  utla_hex_allo %>%
    fortify_hexagon(
      hex_size = hex_size,
      sf_id = "ctyua17cd"
    ) %>%
    left_join(
      weekly_changes %>%
        filter(date == "2021-04-01"),
      by = c("ctyua17cd" = "areaCode")
    ) %>%
    mutate(
      poly_type = "hex",
      group = ctyua17cd
    )

weekly_change_maps_w_legend <-
  plot_map_figure(
    shape_data = weekly_change_map,
    hexagon_data = weekly_change_hexes,
    fill_variable = "weekly_case_change"
  )

figure <-
  plot_grid(
    cp_title(
      "Change in number of COVID-19 cases from the previous week\nin UK Local Authorities on 1 April 2021."
    ),
    weekly_change_maps_w_legend,
    ncol = 1,
    rel_heights = c(0.1, 1)
  )
save_plot(
  "weekly_change_map_figure.pdf",
  figure,
  base_height = 120,
  base_asp = 1.618,
  units = "mm"
)
```

```{r}
coronavirus_data %>%
  ggplot(
    aes(
      x = newCasesBySpecimenDate,
      y = newAdmissions,
      colour = areaCode
    )
  ) +
    geom_point() +
    theme(legend.position = "none") +
    theme_void()
```


```{r}
animated <-
  weekly_changes %>%
  left_join(lad_geo, ., by = c("ctyua17cd" = "areaCode")) %>%
  mutate(log_newCases = log(newCasesBySpecimenDate)) %>%
  ggplot() +
    geom_sf(
      aes(fill = log_newCases),
      lwd = 0.2
    ) +
    theme_map() +
    scale_fill_viridis(discrete = FALSE) +
    # theme(legend.position = "none") +
    transition_manual(date, cumulative = FALSE)

anim_save("map_animation.mp4", animated, nframes = 300, renderer = ffmpeg_renderer())
```

